---
title: "Field descriptions"
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(golfastr)
```

Column-level documentation for the main golfastr datasets. The same tables
are available in R via `pga_field_descriptions()`.

## Leaderboard

Returned by `load_leaderboard()` and `leaderboard()`.

```{r}
knitr::kable(pga_field_descriptions("leaderboard"))
```

## Hole-by-hole

Returned by `load_holes()`.

```{r}
knitr::kable(pga_field_descriptions("holes"))
```

## Strokes gained

Returned by `load_strokes_gained()` and available as the packaged
`strokes_gained` dataset. Values are per-round averages from
ShotLink-measured rounds.

| Field | Description |
|-------|-------------|
| player_id | PGA Tour player ID |
| player_name | Player display name |
| country | Three-letter country code (e.g., "USA", "ENG") |
| sg_putt | Strokes Gained: Putting |
| sg_arg | Strokes Gained: Around the Green |
| sg_app | Strokes Gained: Approach the Green |
| sg_ott | Strokes Gained: Off the Tee |
| sg_t2g | Strokes Gained: Tee to Green |
| sg_total | Strokes Gained: Total |
| rounds | Number of ShotLink-measured rounds |
| season | PGA Tour season year |
