---
title: "Getting started with golfastr"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
```

golfastr provides tidy professional golf data in R: tournament schedules,
leaderboards, hole-by-hole scoring, and strokes gained. Completed PGA Tour
seasons are served from pre-built files hosted on GitHub releases (updated
daily), so loading a full season takes seconds. Everything else falls back
to the live ESPN API.

```{r setup, message = FALSE}
library(golfastr)
library(dplyr)
```

## The season schedule

```{r}
sched <- load_schedule(2025)
head(sched)
```

## Leaderboards

`load_leaderboard()` takes a year and an optional tournament, matched by
event ID or partial name:

```{r}
masters <- load_leaderboard(2025, "Masters")
head(masters)
```

Leave the tournament out to get every completed event in the season in one
tibble:

```{r}
season <- load_leaderboard(2025)
nrow(season)
```

## Hole-by-hole scoring

Hosted seasons include full-field scorecards, one row per player per hole:

```{r}
holes <- load_holes(2025, "Masters")
head(holes)
```

That granularity is enough to rebuild the tournament story. For example,
the cumulative score to par of the top five finishers:

```{r masters-race, message = FALSE}
library(ggplot2)

top5 <- masters |>
  slice_head(n = 5) |>
  pull(player_id)

race <- holes |>
  filter(player_id %in% top5) |>
  arrange(player_id, round, hole) |>
  group_by(player_id, player_name) |>
  mutate(
    hole_index = row_number(),
    to_par = cumsum(score - par)
  ) |>
  ungroup()

ggplot(race, aes(hole_index, to_par, color = player_name)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +
  geom_step(linewidth = 0.8) +
  scale_y_reverse() +
  labs(
    title = "2025 Masters, top five finishers",
    x = "Holes played", y = "Score to par", color = NULL
  ) +
  theme_minimal()
```

## Strokes gained

Per-round strokes gained averages for the current season ship with the
package:

```{r}
load_strokes_gained("Scheffler")
```

## Live data and other tours

Hosted data covers completed PGA Tour events. For tournaments in progress,
pass `live = TRUE` to query ESPN directly:

```{r, eval = FALSE}
load_leaderboard(2026, "Travelers", live = TRUE)
```

LIV Golf, the LPGA, the DP World Tour, and the Champions Tour are available
through the same functions via the `tour` argument:

```{r, eval = FALSE}
load_schedule(2026, tour = "lpga")
load_leaderboard(2026, "Adelaide", tour = "liv")
```

Field-level documentation for every column lives in
`vignette("field-descriptions")` and in `pga_field_descriptions()`.
