---
title: "Plot Golf Data with ggplot2"
description: >
  Load a tournament leaderboard and make a simple scoring chart.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot Golf Data with ggplot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5
)
```

This example shows the leading scores from the 2025 Masters Tournament.
The chart uses hosted data from `golfastr`.

## Load the data

Load `golfastr`, `dplyr`, and `ggplot2`.
Then request the leaderboard by year and tournament name.

```{r load-data}
library(golfastr)
library(dplyr)
library(ggplot2)

masters <- load_leaderboard(2025, "Masters")

leaders <- masters |>
  mutate(score_to_par = as.numeric(score_to_par)) |>
  arrange(score_to_par, total_score) |>
  slice_head(n = 15)

leaders |>
  select(position, player_name, score_to_par) |>
  head()
```

`score_to_par` is the number of strokes above or below par.
A negative value is below par.

## Make the chart

Use `reorder()` to sort the players by score.
Use `coord_flip()` to make the names easy to read.

```{r plot-leaders, fig.alt="Horizontal bar chart of the 15 leading scores at the 2025 Masters Tournament."}
ggplot(
  leaders,
  aes(
    x = reorder(player_name, score_to_par),
    y = score_to_par
  )
) +
  geom_col(fill = "#2d6a4f", width = 0.72) +
  geom_text(
    aes(label = score_to_par),
    hjust = 1.2,
    color = "white",
    size = 3.5
  ) +
  coord_flip() +
  labs(
    title = "2025 Masters Tournament leaders",
    subtitle = "Final score relative to par",
    x = NULL,
    y = "Score relative to par",
    caption = "Data: golfastr"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank())
```

You can use the same code for another tournament.
Change the year and tournament name in `load_leaderboard()`.

